Brief notes on erosion

### 2017-10-19 docs/TexNotes/Notes.tex
Working on chaning multistep to RK2.

% FROM METHODS TEX
\newcommand{\xx}{{\mathbf{x}}}
\renewcommand{\ss}{{\mathbf{s}}}
\newcommand{\nn}{{\mathbf{n}}}
% Derivatives
\newcommand{\pderiv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\tderiv}[2]{\frac{d #1}{d #2}}
% Nick's commands
\newcommand{\thL}{$\theta$--$L$}
\newcommand{\Vn}{V_n}
\newcommand{\Vs}{V_s}
\newcommand{\thalpha}{\pderiv{\theta}{\alpha}}
\newcommand{\elfun}{\zeta}
\newcommand{\thhat}{\hat{\theta}}

@@

Upon transforming to the variables $\theta(\alpha,t)$ and $L(t)$, we obtain the evolution equations
\begin{align}
& \tderiv{L}{t} = - \mean{\thalpha \Vn} \\
& \pderiv{\theta}{t} = \frac{1}{L} \left( \pderiv{\Vn}{\alpha} + \thalpha \Vs \right)
\end{align}
Inserting Eq.~(\ref{eqn:Vn2}) for the normal velocity gives
\begin{equation}
\pderiv{\theta}{t} = \frac{\eps}{L^{2} \log \left(2\pi/L \right)} \ppd{\theta}{\alpha} 
+ \frac{1}{L} \left( \pderiv{\atau}{\alpha} + \thalpha \Vs \right)
\end{equation}
For simplicity, we introduce the variables
\begin{align}
& M = - \mean{\thalpha \Vn} \\
& N = \frac{1}{L} \left( \pderiv{\atau}{\alpha} + \thalpha \Vs \right) \\
& \elfun = \frac{1}{L^{2} \log \left(2\pi/L \right)}
\end{align}
Then the evolution equations can be expressed as
\begin{align}
& \tderiv{L}{t} = M \\
& \pderiv{\theta}{t} = \eps \elfun \ppd{\theta}{\alpha} + N
\end{align}
We treat $\theta$ spectrally
\begin{equation}
\theta(\alpha,t) = \sum_{k = -\infty}^{\infty} \thhat_k e^{2\pi i k \alpha}
\end{equation}
Then we get
\begin{equation}
\tderiv{\thhat_k}{t} +  4 \eps \pi^2 k^2  \elfun \thhat_k = \hat{N}_k
\end{equation}
We introduce the integrating factor
\begin{equation}
\mu = \exp \left( 4 \eps \pi^2 k^2 \int \elfun dt \right)
\end{equation}
so that the solution can be represented as
\begin{align}
\thhat_k(t) & = \frac{1}{\mu} \int^t \mu \hat{N}_k dt'
= \thhat_k(0) + \int_0^t \exp \left( -4 \eps \pi^2 k^2 \int_{t'}^t \elfun dt'' \right) \hat{N}_k \, dt'
% LONGER FORM
%& = \exp \left( -4 \eps \pi^2 k^2 \int_0^t \elfun(L) ds \right) \, 
%\int_0^t \exp \left( 4 \eps \pi^2 k^2 \int_0^s \elfun(L) ds' \right) \hat{N}_k ds \\
\end{align}
where $t'$ and $t''$ are simply dummy variables for integration. We can get the exact relationship
\begin{equation}
\thhat_k(t + \Delta t) = \thhat_k(t) 
+ \int_t^{t+\Delta t} \exp \left( -4 \eps \pi^2 k^2 \int_{t'}^{t+\Delta t} \elfun dt'' \right) \hat{N}_k \, dt'
\end{equation}

>>>>

% FROM METHODS TEX
\newcommand{\xx}{{\mathbf{x}}}
\renewcommand{\ss}{{\mathbf{s}}}
\newcommand{\nn}{{\mathbf{n}}}
% Derivatives
\newcommand{\pderiv}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\tderiv}[2]{\frac{d #1}{d #2}}
% Nick's commands
\newcommand{\thL}{$\theta$--$L$}
\newcommand{\Vn}{V_n}
\newcommand{\Vs}{V_s}
\newcommand{\thalpha}{\pderiv{\theta}{\alpha}}
\newcommand{\elfun}{\zeta}
\newcommand{\thhat}{\hat{\theta}}
\newcommand{\dt}{\Delta t}

@@

Upon transforming to the variables $\theta(\alpha,t)$ and $L(t)$, we obtain the evolution equations
\begin{align}
& \tderiv{L}{t} = - \mean{\thalpha \Vn} \\
& \pderiv{\theta}{t} = \frac{1}{L} \left( \pderiv{\Vn}{\alpha} + \thalpha \Vs \right)
\end{align}
Inserting Eq.~(\ref{eqn:Vn2}) for the normal velocity gives
\begin{equation}
\pderiv{\theta}{t} = \frac{\eps}{L^{2} \log \left(2\pi/L \right)} \ppd{\theta}{\alpha} 
+ \frac{1}{L} \left( \pderiv{\atau}{\alpha} + \thalpha \Vs \right)
\end{equation}
For simplicity, we introduce the variables
\begin{align}
& M = - \mean{\thalpha \Vn} \\
& N = \frac{1}{L} \left( \pderiv{\atau}{\alpha} + \thalpha \Vs \right) \\
& \elfun = \frac{1}{L^{2} \log \left(2\pi/L \right)}
\end{align}
Then the evolution equations can be expressed as
\begin{align}
& \tderiv{L}{t} = M \\
& \pderiv{\theta}{t} = \eps \elfun \ppd{\theta}{\alpha} + N
\end{align}
We treat $\theta$ spectrally
\begin{equation}
\theta(\alpha,t) = \sum_{k = -\infty}^{\infty} \thhat_k e^{2\pi i k \alpha}
\end{equation}
Then we get
\begin{equation}
\label{thODEs}
\tderiv{\thhat_k}{t} +  4 \eps \pi^2 k^2  \elfun \thhat_k = \hat{N}_k
\end{equation}
We introduce the integrating factor
\begin{equation}
\mu_k = \exp \left( 4 \eps \pi^2 k^2 \int \elfun dt \right)
\end{equation}
so that our system becomes
\begin{align}
\label{Leq}
& \tderiv{L}{t} = M \\
\label{theq}
& \tderiv{}{t}\left( \mu_k \theta_k \right) = \mu_k \hat{N}_k
\end{align}
We discretize Eqs.~(\ref{Leq})--(\ref{theq}) with a second-order Runge-Kutta method (the midpoint RK2). In the first step, we go from time $t_n$ and take a half-step to $t_{n+1/2}$
\begin{align}
& L^{(n+1/2)}  = L^{(n)} + \frac{1}{2}\dt \, M^{(n)} \\
& \theta_k^{(n+1/2)} = \nu \left( \theta_k^{(n)} + \frac{1}{2} \dt \,  \hat{N}_k^{(n)} \right)
\end{align}
Here, we have introduced
\begin{equation}
\nu = \frac{\mu_k^{(n)}}{\mu_k^{(n+1/2)}} = \exp \left( -4 \eps \pi^2 k^2 \int_{t_n}^{t_{n+1/2}} \zeta \, dt \right)
\end{equation}
Discretizing with trapezoid rule (and using the fact that $L$ can be solved for first to get $L^{(n+1/2)}$) gives
\begin{equation}
\nu = \exp \left(-2 \eps \pi^2 k^2 \dt \left( \elfun^{(n+1/2)} + \elfun^{n}\right)\right)
\end{equation}
Then we take the second step to $t_{n+1}$ as follows
\begin{align}
& L^{(n+1)}  = L^{(n)} + \dt \, M^{(n+1/2)} \\
& \theta_k^{(n+1)} = \nu \left( \theta_k^{(n)} + \dt \,  \hat{N}_k^{(n+1/2)} \right)
\end{align}
where now we have
\begin{equation}
\nu = \frac{\mu_k^{(n)}}{\mu_k^{(n+1)}} = \exp \left( -4 \eps \pi^2 k^2 \int_{t_n}^{t_{n+1}} \zeta \, dt \right)
\end{equation}
which when discretized becomes
\begin{equation}
\nu = \exp \left(-4 \eps \pi^2 k^2 \dt \left( \elfun^{(n+1)} + \elfun^{n}\right)\right)
\end{equation}
